Analyzing Information Transfer from Different Data Sources via Precise Bias-variance Asymptotics

### 2021-07-07 revision_journal/exp.tex
Updates from Overleaf

%\subsection{Comparison between the averaging and HPS estimators}
\subsection{Regression tasks}

>>>>

\subsection{Comparing HPS to other estimators}

    \begin{align}
    L(\hat{\beta}^{\AV}(\lambda) ) &= \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \lambda^2 \left\| (\Sigma^{(2)})^{1/2}(\beta^{(1)}-\beta^{(2)})\right\|^2 \nonumber \\
    &+  \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \sigma^2   \lambda^2 \bigtr{\Sigma^{(2)} [(X^{(1)})^\top X^{(1)}]^{-1}  }\nonumber\\
    & + \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \sigma^2  (1-\lambda)^2 \bigtr{\Sigma^{(2)} [(X^{(2)})^\top X^{(2)}]^{-1}  }  . \label{L_AVE_simple}
    \end{align}
    and
    \begin{align}
    L(\hat{\beta}_2^{\STL} ) &= \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \sigma^2  \bigtr{\Sigma^{(2)} [(X^{(2)})^\top X^{(2)}]^{-1}  }  . \label{L_STL_simple}
    \end{align}
    %Moreover, there exists a constant $C_0>0$ such that with high probability,
    %\be\label{loss_large}
    %g(a) > g(0) \quad \text{for all $a$ such that }|a|\ge C_0.
    %\ee
    %with high probability for any small constant $\e>0$.

 With Lemma \ref{fact_tr}, we obtain that w.h.p.,
\begin{align}
L(\hat{\beta}_2^{\STL} ) &= \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \frac{p \sigma^2 }{n_2-p}   , \label{L_STL_simple01}
\end{align}
and
\begin{align}
L(\hat{\beta}^{\AV}(\lambda) ) &= \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \lambda^2 \left\| (\Sigma^{(2)})^{1/2}(\beta^{(1)}-\beta^{(2)})\right\|^2 \nonumber \\
&+  \left[ 1+ \OO(p^{-1/2+c})\right] \cdot \sigma^2 \left[\frac{ \lambda^2}{n_1-p}\bigtr{\Sigma^{(2)}(\Sigma^{(1)})^{-1}} + \frac{(1-\lambda)^2p  }{n_2-p}\right] . \label{L_AVE_simple01}
 \end{align}

\todo{merge into previous paragraphs {Simulations.}}
Finally, we validate the result of Example \ref{ex_covshift}.
Figure \ref{fig_covariate} shows task two's prediction loss as we increase task one's sample size.
Recall that $\lambda$ measures the severity of covariate shifts---a larger $\lambda$ means a larger covariate shift.
We indeed observe the dichotomy in Example \ref{ex_covshift} at $n_1 = 800$.
The sample size $n_2$ is $800$ and the noise variance $\sigma^2$ is $1/4$.

%{\cob The idea of proof: proved using free probability; can also use polynomials of random matrices.}

\subsection{Regression tasks}

### 2021-07-07 revision_journal/intro.tex
Updates from Overleaf

\paragraph{Random matrix theory.}
\todo{write a paragraph regarding applications of RMT to modern machine learning.}
\begin{itemize}
	\item \citet{hastie2019surprises}
	\item \citet{bartlett2020benign}
	\item \citet{liang2020just}
	\item \citet{montanari2019generalization}, \citet{liang2020precise}
\end{itemize}

Our proof techniques use the so-called local law of random matrices \cite{erdos2017dynamical}, a recent development in the random matrix theory literature.
In the single-task case, \cite{isotropic} first proved such a local law for sample covariance matrices with isotropic covariance.
\cite{Anisotropic} later extended this result to sample covariance matrices with arbitrary covariance.
%On the other hand, one may derive the asymptotic result in Theorem \ref{thm_main_RMT} with error $\oo(1)$ using the free addition of two independent random matrices in  theory .
These techniques provide almost sharp convergence rates to the asymptotic limit compared to other methods such as free probability \cite{nica2006lectures}.
To the best of our knowledge, we are not aware of any previous results in the multi-task case, even for two tasks (with arbitrary covariate shifts).
%\HZ{can you please add a discussion of the related work from RMT?}

>>>>

\paragraph{Random matrix theory.}
\todo{write a paragraph regarding applications of RMT to modern machine learning.}
\begin{itemize}
	\item \citet{hastie2019surprises}
	\item \citet{bartlett2020benign}
	\item \citet{liang2020just}
	\item \citet{montanari2019generalization}, \citet{liang2020precise}
\end{itemize}
\begin{remark}
    \todo{clarify overparametrize}
    As a clar,  following the work of \citet{hastie2019surprises} (see also \citet{bartlett2020benign} and the references therein)
\end{remark}

These previous works mainly consider cases where there is only one high-dimensional sample covariance matrix involved. %need to check 
In the single task case, if the data is Gaussian, then the precise variance limit can be derived directly from properties of Wishart distribution. In the high-dimensional setting, the eigenvalues of Wishart matrices satisfy the famous Marchenko–Pastur (MP) law \cite{MP}, whose Stieltjes transform characterizes the variance limit. Furthermore, it is well-known that the MP law holds universally regardless of the detailed data distribution \cite{bai2010spectral}. \cite{isotropic} obtained a sharp converge rate of the empirical spectral distribution (ESD) to the MP law for sample covariance matrices with isotropic covariance. \cite{Anisotropic,DY} later extended this result to sample covariance matrices with arbitrary population covariances. These results are proved by establishing certain optimal convergence estimates of the Stieltjes transforms of sample covariance matrices, which are referred to as \emph{local laws} in the random matrix theory literature. We refer the reader to \cite{erdos2017dynamical} and the references therein for a detailed review of related concepts. In this paper, we extend these techniques to the two-task setting and prove the local law for the sum of two sample covariance matrices with an arbitrary covariate shift, which is one technical contribution of this work. From the local law, we can derive the variance limit with a precise dependence on the singular values of the covariate shift matrix.

%Our proof techniques use the so-called local law of random matrices, a recent development in the random matrix theory literature, called the \emph{local law} of

%These techniques provide almost sharp convergence rates to the asymptotic limit compared to other methods such as free probability \cite{nica2006lectures}.
%We also extend the techniques in ?, which is one contribution of this work.

%On the other hand, one may derive the asymptotic result in Theorem \ref{thm_main_RMT} with error $\oo(1)$ using the free addition of two independent random matrices in theory.

There is another method to derive the asymptotic variance limit using free addition in free probability theory \cite{nica2006lectures}. We do not take this procedure because this method is not completely justified in the general non-Gaussian setting with non-diagonal covariate shift matrix. 
%(in the sense that it is hard to establish rigorously the free independence between two sample covariance matrices). 
Furthermore, our techniques provide almost sharp convergence rates to the asymptotic limit, which is not given by free probability techniques.

On the other hand, the bias limit is much harder to derive. However, when the data is Gaussian distributed and there is no covariante shift, we can use recent results in ... to derive the bias limit. To the best of our knowledge, we are not aware of any previous results in the multi-task case, even for cases without covariate shifts.

%\HZ{can you please add a discussion of the related work from RMT?}
